SIMD\slash SIMT-vectorized Thermo-chemical Source Term Evaluation

### 2018-02-14 paper.tex
updates

\subsection{Code Generation and Testing Infrastructure}
Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL\slash C-code.
This allows for easy changes of program structure, e.g. data ordering, vectorization, threading patterns and easy switching of target-languages (and simpler extension to additional languages, e.g., CUDA).
Additionally, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, and via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the autodifferentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian subcomponents.
Validation of the complete source-terms and Jacobian codes will be discussed in detail in~\cref{s:validation}.

@@

This point is further underscored by the error in the net ROP, which is \textasciitilde\numrange{3}{9} (or \numrange{7}{9} for the P-Log excluded error) orders of magnitude larger than the error in either the forward or reverse ROP.
In~\cref{T:source_error}, we see that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increases as the error in species rates is amplified by multiplication with thermodynamic properties.
Again, we note that the above discussion does not imply that discrepencies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.

@@

%16.1.1 & 1.0 \slash 5.0.0 \\

\begin{table}[htb]
\centering
\begin{tabular}{@{}l l l l l@{}}
\toprule
Model & Instruction Set & Vector Width & Cores & Identifier \\
\midrule
Intel Xeon X5650 & SSE4.2 & \SI{128}{\bit} & $2 \times 6$ & \texttt{sse4.2} \\
Intel E5-2690 V3 & AVX2 & \SI{256}{\bit} & $2 \times 12$ & \texttt{avx2} \\
\bottomrule
\end{tabular}
\label{t:cpus}
\caption{The CPU configurations used in this study.}
\end{table}
\begin{table}
\centering
\begin{tabular}{@{}l l l l l@{}}
\toprule
Model & CUDA Version & CUDA Cores & Identifier \\
\midrule
Tesla C2075 & 9.0.176 & \num{448} & \texttt{C2075} \\
Tesla K40m & 8.0.61 & 2880 & \texttt{K40m} \\
\bottomrule
\end{tabular}
\label{t:gpus}
\caption{The GPU configurations used in this study.}
\end{table}

>>>>

\subsection{Code Generation and Testing Infrastructure}
Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL\slash C-code.
This allows for easy changes of program structure, e.g. data ordering, vectorization, threading patterns and easy switching of target-languages (and simpler extension to additional languages, e.g., CUDA).
Additionally, \texttt{loo.py} can execute developed subroutines from python (natively for C-code, and via \texttt{PyOpenCL}~\cite{kloeckner_pycuda_2012} for OpenCL-code), enabling unit-testing\slash validation for each component of the Jacobian or source terms.
The source-terms and subcomponents thereof (e.g., rates of progress, pressure-modification terms, etc.) are directly compared to Cantera~\cite{Cantera}, while the automatic differentiation code Adept~\cite{adept-v11} is used to obtain reference answers for Jacobian subcomponents.
Validation of the complete source-terms and Jacobian codes will be discussed in detail in~\cref{s:validation}.

@@

This point is further underscored by the error in the net ROP, which is \textasciitilde\numrange{3}{9} orders of magnitude (or \numrange{7}{9} orders of magnitude for the P-Log excluded error) larger than the error in either the forward or reverse ROP.
In~\cref{T:source_error}, we see that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
The molar species production rate error norm is similar in magnitude to that of the net ROP, however the temperature and state parameter rate errors again increases as the error in species rates is amplified by multiplication with thermodynamic properties.
Again, we note that the above discussion does not imply that discrepencies in computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.

\subsection{Jacobian validation}

As in our previous work~\cite{Niemeyer:2016aa}, the correctness of the Jacobian codes generated by \texttt{pyJac} was determined by comparison to Jacobians obtained by automatic differentiation of the source rate subroutines via the \texttt{Adept} software library~\cite{adept-v11,hogan2014fast}.
The reasons for this choice are explained fully in~\cite{Niemeyer:2016aa}, but broadly speaking automatic differentiation provides relatively fast, highly accurate Jacobian matrix evaluation with relatively programming little effort.
The discrepency between the analytical and automatic differentiated Jacobians for thermochemical state $k$, denoted by $\mathcal{J}_k$ and $\hat{\mathcal{J}}_k$ respectively, is determined by the relative error Frobenius norm over all Jacobian indicies $i, j$:
\begin{equation}
 \label{e:jac_error_base}
 E_{\text{rel}, k} = \frac{\left\lVert \hat{\mathcal{J}}_{ijk} - \mathcal{J}_{ijk} \right\rVert_{i,j,F}}{\left\lVert \hat{\mathcal{J}}_{ijk} \right\rVert_{i,j,F}}
\end{equation}

To avoid large relative errors of small nonzero Jacobian elements due to accumulation of floating point error, the Frobenius norm of the automatically differentiated Jacobian is calculated over all thermochemical states $k$:
\begin{equation}
 \label{e:thresh}
 \mathcal{T} = \left\lVert \mathcal{\hat{J}} \right\rVert_{k, F}
\end{equation}
The error statistics reported in this section are then based only on matrix elements where $\mathcal{J}_{ijk} \ge \frac{\mathcal{T}}{\mathcal{C}}$; this filtered form of~\cref{e:jac_error_base} is denoted $E_{\mathcal{C}}$.

\todo{rerun, especially H2 error}
\begin{table}[tbp]
\centering
\begin{tabular}{@{}l S S[table-format=1.3e1] S[table-format=1.3e1] S[table-format=1.3e1]@{}}
\toprule
Model                 & \multicolumn{1}{c}{Sample size}   & \multicolumn{1}{c}{$\bar{\mathcal{T}}$}  & \multicolumn{1}{c}{$E_{10^{20}}$}   & \multicolumn{1}{c}{$E_{10^{15}}$} \\
\midrule
\ce{H2}\slash \ce{CO} & \num{900900}                      & \num{6.431e+18}      & \num{8.083e-1}  & \num{1.907e-5} \\
GRI-Mech 3.0          & \num{450900}                      & \num{7.783e+19}      & \num{1.890e-7}  & \num{1.877e-7} \\
USC-Mech II           & \num{91800}                       & \num{2.830e+21}      & \num{5.567e-3}  & \num{1.704e-7} \\
\ce{iC5H11OH}         & \num{450900}                      & \num{2.733e+26}      & \num{1.136e-3}  & \num{2.764e-5} \\
\bottomrule
\end{tabular}
\caption{Summary of Jacobian matrix validation results.
Error statistics are the largest filtered relative error $E_\mathcal{C}$ for over all samples.
It is noted that the threshold described in~\cref{e:thresh} varies slightly between the \conp~and \conv~cases; the reported $\bar{\mathcal{T}}$ is the average of the two, however the appropriate value was used during calculations of the error statistics..
}
\label{T:error}
\end{table}

\Cref{T:error} presents the error statistics of \texttt{pyJac} over the tested PaSR thermochemical condition database.
For all tested cases, \todo{need srv2 nvidia tests to really confirm here} the maximum filtered error norm is less than unity.
The filtered error norm is largest for the \ce{H2}\slash\ce{CO} model with the more stringent tolerance ($\mathcal{C} = 10^{20}$).
We note that $\mathcal{T}$ for this model is smaller than the tolerance in this case, and hence we are comparing numbers smaller than $\mathcal{O}(1)$; relaxing the tolerance to $\mathcal{C} = 10^{15}$ significantly lowers the filtered error norm to $\mathcal{O}(10^{-5})$.
Similar error characteristics are seen for all models except GRI-Mech 3.0---in which the error norms for both tolerances are $\sim\mathcal{O}(10^{-7})$---that is a larger error norm for the more stringent tolerance that drops 2--4 orders of magnitude for the relaxed tolerance.

@@

\begin{table}[htb]
\centering
\begin{tabular}{@{}l l l l l@{}}
\toprule
Model & Instruction Set & Vector Width & Cores & Identifier \\
\midrule
Intel Xeon X5650 & SSE4.2 & \SI{128}{\bit} & $2 \times 6$ & \texttt{sse4.2} \\
Intel E5-2690 V3 & AVX2 & \SI{256}{\bit} & $2 \times 12$ & \texttt{avx2} \\
\bottomrule
\end{tabular}
\label{t:cpus}
\caption{The CPU configurations used in this study.}
\end{table}
\begin{table}
\centering
\begin{tabular}{@{}l l l l l@{}}
\toprule
Model & NVIDIA Driver Version & CUDA Cores & Identifier \\
\midrule
Tesla C2075 & 384.81 & \num{448} & \texttt{C2075} \\
Tesla K40m & 375.26 & \num{2880} & \texttt{K40m} \\
\bottomrule
\end{tabular}
\label{t:gpus}
\caption{The GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.}
\end{table}
